---
title: "MassNBCtools-NEWS"
author: "Ben.Block@tetratech.com"
date: "`r Sys.time()`"
output: github_document
  #md_document:
   # variant: markdown_github
---

<!-- NEWS.md is generated from NEWS.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "NEWS-"
)
```

```{r, echo = FALSE}
cat(paste0("Last Update: ",Sys.time()))
```

# AKDECtools 1.0.1.9001 (2024-04-10)
-   Update:
    - Aesthetic updates to boxPlot and timeSeries functions
		- Added filter of NA constituents to MagDurFreq_combine function
		- Added if no samples output to MagDurFreq_hardness function
		- Added fraction consideration to MagDurFreq and MagDurFreq_pH functions
		- Renamed categorize_AU function to categorize_AU_uses
		- Added simplify standards argument to categorize_AU_uses function
		- Added categorize_AU function that only categorizes the entire AU
    - Updated vignette
    
# AKDECtools 1.0.1.9001 (2024-04-10)
-   Update: 
    - Added Open Sans font to boxPlot and timeSeries functions
    
# AKDECtools 1.0.0 (2024-02-27)
-   Version 1.0.0 release
-   Update: 
    - Rectified outstanding function issues
    - Finalized documentation of functions and vignette
    - Added two helper functions

# AKDECtools 0.0.2.9000 (2024-02-21)

-   Update: 
    - Rectified outstanding function issues
    - Nearly finalized documentation of functions and vignette
    
# AKDECtools 0.0.1.9100 (2024-02-12)

-   Update: 
    - Updated plotting functions
    - Added Magnitude, frequency, duration functions
    
# AKDECtools 0.0.1.9040 (2024-01-05)

-   Update: 
    - Updated function documentation
    - Added vignette
    
# AKDECtools 0.0.1.9021 (2024-01-05)

-   Update: Added multiple functions
    -   simplePull
    -   boxPlot
    -   boxPlotlog10
    -   timeSeries
    -   timeSerieslog10

# AKDECtools 0.0.0.9000 (2023-10-18)

-   Initial development
