---
title: "AKDECtools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AKDECtools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(AKDECtools)
```

## Introduction

The following vignette is associated with the [*AKDECtools*](https://github.com/Blocktt/AKDECtools) R package developed by Tetra Tech for the Alaska Department of Environmental Conservation (AK DEC). This vignette assumes that the user has relatively recent versions of [R](https://www.r-project.org/) and [RStudio](https://posit.co/download/rstudio-desktop/) installed and that they are familiar with the R coding language. Please direct any questions regarding the usage of the *AKDECtools* R package to Amber Bethe, AK DEC (amber.bethe\@alaska.gov).

### Purpose

The *AKDECtools* package has many functions that are useful to AK DEC staff and may be useful to others outside of that organization. Some functions are designed to manipulate data, others to evaluate water quality standards, and some are used for data visualization. Simply run `help(package = "AKDECtools")` to access the list of the package's functions. Each function is explained in detail below with examples.

## Installing AKDECtools

*AKDECtools* can be downloaded from Ben Block's GitHub repository using the following R code. Note, if you want to download this vignette with the package, use the second function that includes the vignette argument.

```{r, eval = FALSE}
# install.packages("devtools") # install devtools if necessary
devtools::install_github("Blocktt/AKDECtools")

# run the following to download vignette with the package
devtools::install_github("Blocktt/AKDECtools", build_vignettes = TRUE)
```

# Data manipulation functions

## simplePull

The *simplePull* function is intended to filter specific data from a larger WQ dataset by either an assessment unit identifier (AU_ID), water quality constituent (TADA.CharacteristicName), or both. See function documentation by running `?AKDECtools::simplePull`. Note: TADA.CharacteristicName comes from the *TADA* R package (for AK DEC staff users, run raw WQP data through [*data_processing.R*](https://github.com/KateriSalk/Alaska_IR_Automation/blob/main/Code/5_Data_Processing/data_processing.R)) before running this function.

Below are examples on how to use *simplePull* depending on whether a user wants to filter the input dataset by one or more AU_ID(s) and/or one or more constituent(s).

```{r, eval=FALSE}
# Load packages
library(readr) # readr is a tidyverse package to read CSVs
library(AKDECtools)

# Import the example file from the AKDECtools package
df_ExampSamps <- read_csv(system.file("extdat/AK_Example_Samples.csv", package = "AKDECtools"), guess_max = 10^6)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Example 1 - single AU and single constituent

test_pull <- simplePull(data = df_ExampSamps, AU_ID = c('AK_R_1010504_005')
, constituent = c('PH'))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Example 2 - multiple AUs and single constituent

test_pull <- simplePull(data = df_ExampSamps, AU_ID = c('AK_R_1010504_005', 'AK_B_1010203_001')
, constituent = c('TEMPERATURE, WATER'))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Example 3 - single AU and multiple constituents

test_pull <- simplePull(data = df_ExampSamps, AU_ID = c('AK_R_1010504_005')
, constituent = c('PH', 'TEMPERATURE, WATER'))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Example 4 - multiple AUs and multiple constituents

test_pull <- simplePull(data = df_ExampSamps, AU_ID = c('AK_R_1010504_005', 'AK_B_1010203_001')
, constituent = c('PH', 'TEMPERATURE, WATER'))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Evaluate unique results in test_pull using the following code
unique(test_pull$AU_ID) # returns unique AU_IDs in trimmed dataset
unique(test_pull$TADA.CharacteristicName) # returns unqiue TADA.CharacteristicNames in trimmed dataset
```

# Water quality standards functions

# Graphing functions

## boxPlot

See function documentation by running `?AKDECtools::boxPlot`.

## boxPlotlog10

See function documentation by running `?AKDECtools::boxPlotlog10`.

## timeSeries

See function documentation by running `?AKDECtools::timeSeries`.

## timeSerieslog10

See function documentation by running `?AKDECtools::timeSerieslog10`.
